Reproducible Research
========================================

```{r setOptions, echo = FALSE}
#install.packages("knitr") #Install the package if you do not have one.
library(knitr)
opts_chunk$set(echo = FALSE)
```

###Q0. Loading and Preprocessing the data
```{r, echo = TRUE, cache= TRUE}
activity <- read.csv("activity.csv", sep = ",")
```

###Q1. What is mean total number of steps taken per day?
```{r totalstep.plot, echo = TRUE, results="asis"}
#Split the dataframe by the date. Assuming that the days are continuous
by.day <- split(activity, activity$date)
total.step.day <- lapply(X = by.day, FUN = function(x) {sum(x$steps)})
#Makes the dataframe with the date and its total step
data.frame.step <- data.frame(Date = as.Date(names(total.step.day)), total = NA)
for(i in 1:nrow(data.frame.step)) {
  data.frame.step[i, 2] <- total.step.day[[i]]
}
#Plots the day and total step
plot(data.frame.step$Date, data.frame.step$total, type ="h", 
     lwd = 5, col = "black",
     xlab = "Date", ylab = "Total # of Step",
     main = "Total number of steps per day from 10-01 to 11-30")
```

The Code Below Calcuates the mean and median of the total.steps.  
It appears that mean and median are similar.
```{r mean.median, echo = TRUE} 
mean.step <- mean(data.frame.step$total, na.rm = TRUE)
median.step <- median(data.frame.step$total, na.rm = TRUE)
```

####The mean total step is `r mean.step`   
####The median total step is `r median.step`  

<br>

###Q2. What is the average daily activity pattern?

```{r, echo = TRUE}
#install.packages("reshape2") #install the package if you do not have the package
library(reshape2)
#Process the original data so that it is appropriate for this question.
transform.melt <- melt(activity, id.vars = "interval", measure.vars = "steps", na.rm = TRUE)
transform.dcast <- dcast(transform.melt, formula = interval ~ variable, mean)
#Plots the data in time series
plot(transform.dcast$interval, transform.dcast$steps, type = "l", 
     xlab = "Interval (5 minutes)", ylab = "Average steps",
     main = "Average Daily Pattern", lwd = 1.5)
```

```{r max, echo = TRUE}

maximum.steps <- max(transform.dcast$steps)
maximum.index <- grep(maximum.steps, transform.dcast$steps)
maximum.interval <- transform.dcast[maximum.index, 1]

```

####The maximum interval is `r maximum.interval`  
####Its value is `r maximum.steps`  

<br>

###Q3. Imputing missing values

```{r numofNA, echo = TRUE}
not.available <- is.na(activity$steps)
#Number of NA's in the steps column of the dataset. True if NA
table(not.available)
```

```{r print, echo=FALSE}
nas <- sum(is.na(activity$steps))
print(paste("There are total of", nas, "NAs"))
```

#### I decided to impute the missing value using 5 minute interval.

```{r imputeNa, echo=TRUE}
by.day.impute <- by.day
for(i in 1:length(total.step.day)) {
  if (is.na(total.step.day[[i]])) {
    by.day.impute[[names(total.step.day[i])]]$steps <- transform.dcast[2]
  }
}
total.step.day.impute <- lapply(X = by.day.impute, FUN = function(x) {sum(x$steps)})
#Makes the dataframe with the date and its total step
data.frame.step.impute <- data.frame(Date = as.Date(names(total.step.day.impute)), total = NA)
for(i in 1:nrow(data.frame.step)) {
  data.frame.step.impute[i, 2] <- total.step.day.impute[[i]]
}

#Plots the day and total step
plot(data.frame.step.impute$Date, data.frame.step.impute$total, type = "h",
     lwd = 5, col = "black",
     xlab = "Date", ylab = "Total # of Step",
     main = "Total # of steps per day from 10-01 to 11-30 (Imputed NA Values)")
```

```{r imputeMean, echo=TRUE} 
impute.mean <- mean(data.frame.step.impute$total)
impute.median <- median(data.frame.step.impute$total)

difference.mean <- paste(((impute.mean - mean.step) / mean.step) * 100, "%")
difference.median <- paste(((impute.median - median.step) / median.step) * 100, "%") 
```
####The mean of imputed missing value is `r impute.mean`  
####The median of imputed missing value is `r impute.median`  
####Difference in mean value from original data and imputed data is `r difference.mean`  
####Difference in median value from original data and imputed data is `r difference.median`
####From the difference, we barely see difference in mean and median between original data and imputed missing value data. But we observe that imputed missing value graph does not have any gaps in the bin (Because all the values for each day are filled in with some data). 

<br>

###Q4. Are there differences in activity patterns between weekdays and weekends?



