Reproducible Research
========================================

```{r setOptions, echo = FALSE}
library(knitr)
opts_chunk$set(echo = FALSE)
```

###Q0. Loading and Preprocessing the data
```{r, echo = TRUE, cache= TRUE}
activity <- read.csv("activity.csv", sep = ",")
```

###Q1. What is mean total number of steps taken per day?
```{r totalstep.plot, echo = TRUE, results="asis"}
#Split the dataframe by the date. Assuming that the days are continuous
by.day <- split(activity, activity$date)
total.step.day <- lapply(X = by.day, FUN = function(x) {sum(x$steps)})
#Makes the dataframe with the date and its total step
data.frame.step <- data.frame(Date = as.Date(names(total.step.day)), total = NA)
for(i in 1:nrow(data.frame.step)) {
  data.frame.step[i, 2] <- total.step.day[[i]]
}
#Plots the day and total step
plot(data.frame.step$Date, data.frame.step$total, type = "h",
     lwd = 5, col = "black",
     xlab = "Date", ylab = "Total # of Step",
     main = "Total number of steps per day from 10-01 to 11-30")
```

The Code Below Calcuates the mean and median of the total.steps.  
It appears that mean and median are similar.
```{r mean.median, echo = TRUE} 
mean.step <- mean(data.frame.step$total, na.rm = TRUE)
median.step <- median(data.frame.step$total, na.rm = TRUE)
```

####The mean total step is `r mean.step`   
####The median total step is `r median.step`  

<br>

###Q2. What is the average daily activity pattern?

```{r, echo = TRUE}
library(reshape2)
#Process the original data so that it is appropriate for this question.
transform.melt <- melt(activity, id.vars = "interval", measure.vars = "steps", na.rm = TRUE)
transform.dcast <- dcast(transform.melt, formula = interval ~ variable, mean)
#Plots the data in time series
plot(transform.dcast$interval, transform.dcast$steps, type = "l", 
     xlab = "Interval (5 minutes)", ylab = "Average steps",
     main = "Average Daily Pattern", lwd = 1.5)
```

```{r max, echo = TRUE}

maximum.steps <- max(transform.dcast$steps)
maximum.index <- grep(maximum.steps, transform.dcast$steps)
maximum.interval <- transform.dcast[maximum.index, 1]

```

####The maximum interval is `r maximum.interval`  
####Its value is `r maximum.steps`  

<br>

###Q3. Imputing missing values

```{r numofNA, echo = TRUE}
not.available <- is.na(activity$steps)
table(not.available)
```

```{r print, echo=FALSE}
nas <- sum(is.na(activity$steps))
print(paste("There are total of", nas, "NAs"))
```




