---
title: "Statistical Inference"
author: "Jaewon Chang"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#### Purpose: This report is based on the investigation of forty different exponential distributions in R software to compare it with the CLT (Central Limited Theorem). I used a specific parameter for this project, where lambda equals 0.2.


```{r part.one}
#For Using a randomly generated values
set.seed(1)

#Given lambda value for exponential distribution
lambda <- 0.2
#Given number of exponential values
n.exp <- 40

#Number of simulation for CLT
n.simulation <- 1000

#Random exponential simulation for one thousand times.
random.exp <- replicate(n = n.simulation, expr = rexp(n = n.exp, rate = lambda))
mean.matrix <- apply(random.exp, 2, mean)

#Statistical values (empirical values)
mean.value <- mean(mean.matrix)
sd.value <- sd(mean.matrix)
var.value <- sd.value^2

#Theoretical Values
theo.mean <- 1 / lambda
theo.sd <- theo.mean / sqrt(n.exp)
theo.var <- theo.sd^2

mean <- c(theo.mean, mean.value)
sd <- c(theo.sd, sd.value)
var <- c(theo.var, var.value)

#Matrix that shows the statistical values of theoretical value and empirical vlaue of randomly generated exponential distribution.

#Mean and variances of theoretical and empirical values are quite close to each other.
statistics <- rbind(mean, sd, var)
colnames(statistics) <- c("Theoretical", "Empirical")
statistics

#Graph
hist(mean.matrix, breaks = 40, main = "Compare to Normal Distribution", col = "salmon", xlab = "Mean")

dnorm.seq <- seq(min(mean.matrix), max(mean.matrix), length = 100)
dnorm.y <- dnorm(x = dnorm.seq, mean = theo.mean, sd = theo.sd)
lines(dnorm.seq, dnorm.y * 100)
```

Based on the comparison table (Theoretical and Empirical) above and comparison graph above, they appear to be very close to each other; the frequency bars falls under the normally distributed lines and theo/empirical valuese are quite similar.

